---
title: How Far in Advance Users Schedule Posts
author: ''
date: '2018-12-03'
slug: how-far-in-advance-users-schedule-posts
categories: []
tags: []
---

Buffer is a tool that allows people to schedule their social media posts in advance. In this analysis we will explore how far in advance users schedule their posts. If users don't schedule many posts very far out in advance, it may have some implications for how we develop the product in the future.

To do this we will compute some summary statistics for a sample of users that have scheduled posts in the past six months. Then, we will summarise those summary statistics to try to understand the overall population of users.

## Summary
Almost all users scheduled most (at least 75%) of their updates to be sent within 10 days. Around 53% of users schedule most of their posts to be sent within 24 hours. This means that around 47% of users scheduled most of their posts to be sent more than 24 hours later. 

Around 67% of users scheduled most of their posts to be sent within 48 hours, which implies that a third of users scheduled most of their posts to be sent more than 48 hours later.

Around 94% of users scheduled most of their posts to be sent within the week, which implies that around 6% of users scheduled most of their posts to be sent more than one week later.

## Data Collection
We'll gather one million updates that were created in the past month.

```{r include = FALSE, message = FALSE, warning = FALSE}
# load packages
library(buffer)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
```

```{r include = FALSE, eval = FALSE}
# connect to redshift
con <- redshift_connect()
```

```{sql eval = FALSE}
select 
  up.id as update_id
  , up.user_id as user_id
  , u.billing_plan_name as plan
  , up.created_at
  , up.sent_at
  , datediff(hour, up.created_at, up.sent_at) as hours_to_send
  , up.profile_service
from dbt.updates as up
inner join dbt.users as u
  on up.user_id = u.id
where up.created_at > current_date - 180
and up.was_sent_with_buffer
and up.created_at < current_date - 60
order by random()
limit 1000000
```

```{r include = FALSE, eval = FALSE}
# set as integer
updates$hours_to_send <- as.integer(updates$hours_to_send)

# remove nas
updates <- updates %>% filter(!is.na(hours_to_send) & hours_to_send >= 0)

# save data
saveRDS(updates, file = 'time_to_send_updates.rds')
```

```{r include = FALSE}
# load data
updates <- readRDS('time_to_send_updates.rds')
```

We have around 918 thousand updates created in the past six months. 

## Exploratory Plots
Let's begin by plotting the distribution of the number of hours in advance people scheduled their posts. The actual metric that we are plotting the distribution of is the difference between the updates' `created_at` and `sent_at` times. The difference is measured in hours.

```{r echo = FALSE, warning = FALSE, message = FALSE}
# plot distribution of hours to send
ggplot(updates, aes(x = hours_to_send)) +
  geom_density(fill = "black", alpha = 0.4) +
  scale_x_continuous(limits = c(0, 240)) +
  buffer_theme() +
  labs(x = "Number of Hours Between Created and Sent Dates", y = NULL,
       title = "Number of Hours Scheduled in Advance",
       subtitle = "All posts created in past six months")
```

As we might have guessed, most posts are scheduled to go out in a very short period of time and there is a long tail of updates that were scheduled far in advance. This distribution visualizes the differences between `created_at` and `sent_at` times for all updates. We're more interested in looking at this data on a _user_ basis, so let's create a summary statistic for each user.

We'll use the 75th percentile for the `hours_to_send` metric all users that have scheduled at least 4 updates. This will tell us, for each user, the number of hours for which 75% of their posts were scheduled in advance.

```{r}
# calculate 75% quantile
by_user <- updates %>% 
  group_by(user_id) %>% 
  summarise(updates = n_distinct(update_id),
            percentile = quantile(hours_to_send, probs = 0.75)[[1]]) %>% 
  filter(updates >= 4)
```

Now let's plot this metric for the 50 thousand users in our dataset.

```{r echo = FALSE, message = FALSE, warning = FALSE}
# plot distribution of percentiles
ggplot(by_user, aes(x = percentile)) +
  geom_density(fill = "black", alpha = 0.4) +
  scale_x_continuous(limits = c(0, 240), breaks = seq(0, 240, 24)) +
  scale_y_continuous(labels = percent) +
  buffer_theme() +
  labs(x = "Number of Hours Between Created and Sent Dates", y = NULL,
       title = "Number of Hours Scheduled in Advance",
       subtitle = "Number of Hours For Which 75% of Posts Are Scheduled In Advance")
```

Most users schedule most of their updates to be sent within 24 hours. However, there is also a long tail of users who have scheduled most of their updates to be sent far into the future. Visualizing the cumulative distribution function (CDF) will help us better summarise the effects of this long tail.

```{r echo = FALSE, message = FALSE, warning = FALSE}
# plot distribution of percentiles
ggplot(by_user, aes(x = percentile)) +
  stat_ecdf() +
  coord_cartesian(xlim = c(0, 240)) +
  scale_x_continuous(limits = c(0, 240), breaks = seq(0, 240, 24)) +
  scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.2)) +
  buffer_theme() +
  labs(x = "Number of Hours Between Created and Sent Dates", y = NULL,
       title = "Number of Hours Scheduled in Advance",
       subtitle = "Number of Hours For Which 75% of Posts Are Scheduled In Advance")
```

The CDF shows us that almost all users schedule most of their updates to be sent within 10 days. Around 53% of users schedule most of their posts to be sent within 24 hours. This means that around 47% of users scheduled most of their posts to be sent more than 24 hours later. 

Around 67% of users scheduled most of their posts to be sent within 48 hours, which implies that a third of users scheduled most of their posts to be sent more than 48 hours later.

Around 94% of users scheduled most of their posts to be sent within the week, which implies that around 6% of users scheduled most of their posts to be sent more than one week later.
