---
title: Where should the next Buffer retreat be?
author: ''
date: '2019-04-04'
slug: where-should-the-next-buffer-retreat-be
categories: []
tags: []
---

Buffer is a fully remote and distributed team with over 80 team members based around the world. Once a year, we all gather for a company retreat. This year, the retreat is in San Diego. After seeing [this great blog post](https://livefreeordichotomize.com/2018/06/27/bringing-the-family-together-finding-the-center-of-geographic-points-in-r/), I thought it would be fun to identify the geographic midpoint of all of Buffer's team members.

```{r warning = FALSE, message = FALSE}
# load packages 
library(tidyverse)
library(leaflet)
library(opencage)
library(geosphere)
library(RColorBrewer)
```

## Locating Team Members
The location of every team member can be found in [our transparent salary spreadsheet](https://docs.google.com/spreadsheets/d/11s9VSyf4yaYUsqBKLaVH78NL8wdl8gXoj5BGAzjIFuc/edit#gid=671465451). We'll import the locations from a CSV.

```{r eval = FALSE}
# read data from csv
locations <- read.csv("~/Downloads/locations.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r include = FALSE}
#saveRDS(locations, file = "team_locations.rds")

locations <- readRDS("team_locations.rds")

# fix Victoria BC
locations$location <- gsub("Victoria, B.C.", "Victoria, Canada", locations$location)
locations$location <- gsub("Pleasanton, CA, USA", "Oakland, CA, USA", locations$location)
locations$location <- gsub("Singapore", "Singapore, Singapore", locations$location)
```

This is what the data looks like.

```{r echo = FALSE}
DT::datatable(locations)
```

To get the longitude and latitude for each location, we'll use the`opencage` package. The `open_forward()` function will give multiple results for each call, along with a confidence score for each. We can sort by the confidence score and take the top result.

```{r}
# function to get most likely latitude and longitude
geocode_locations <- function(location) {
  
  opencage_forward(location)$results %>%
    arrange(desc(confidence)) %>%
    select(city = query, lat = geometry.lat, lon = geometry.lng) %>%
    slice(1)
}
```

```{r}
# group by city 
by_city <- locations %>% 
  group_by(location) %>% 
  summarise(number = n())

# map locations
geocodes <- map_df(by_city$location, geocode_locations)

# join to original data frame
by_city <- by_city %>% 
  inner_join(geocodes, by = c("location" = "city")) %>% 
  filter(location != "Nomad")

# join to locations data frame too
locations <- locations %>% 
  inner_join(geocodes, by = c("location" = "city")) %>% 
  filter(location != "Nomad")
```

Now we can plot all of the locations.

```{r echo = FALSE}
# color palette
cols <- paletteer::paletteer_d("LaCroixColoR", 
                    "PassionFruit", 
                    n = 12, 
                    type = "continuous")

leaflet() %>%
  addTiles() %>%  
  addCircleMarkers(lng = locations$lon,
                   lat = locations$lat,
                   label = locations$name,
                   color = cols,
                   radius = 6)
```

## Geographic Average
The easiest way is to find the geographic center is to average all of the longitude points and average all of the latitude points. This can be weighted by the number of people at each point by taking a weighted mean. 

```{r}
# average latitude and longitude
geographic_average <- function(lon, lat, weight = NULL) {
  
  if (is.null(weight)) {
    weight <- rep(1, length(lon))
  }
  
  lon <- weighted.mean(lon, w = weight)
  lat <- weighted.mean(lat, w = weight)
  data.frame(lon = lon, lat = lat)
}
```

## Geographic Midpoint
A bit of a more complex version uses vector algebra – first converting the points to radians, then sticking them on a Cartesian plane. This method is well explained on [Ask Dr. Math](http://mathforum.org/library/drmath/view/68373.html).

```{r}
# function to find geographic midpoint
geographic_midpoint <- function(lon, lat, weight = NULL) {
  
  if (is.null(weight)) {
    weight <- rep(1, length(lon))
  }
  
  # degrees to radians
  lat <- lat * pi / 180
  lon <- lon * pi / 180
  
  # cartesian coordinates
  x <- cos(lat) * cos(lon)
  y <- cos(lat) * sin(lon)
  z <- sin(lat)
  
  # weighted mean
  x <- weighted.mean(x, w = weight)
  y <- weighted.mean(y, w = weight)
  z <- weighted.mean(z, w = weight)
  
  # convert to lat and lon
  lon <- atan2(y, x) * 180 / pi
  hyp <- sqrt(x * x + y * y)
  lat <- atan2(z, hyp) * 180 / pi
  
  data.frame(lon = lon, lat = lat)
}
```

Let's calculate these points and map them out.

```{r}
df <- by_city %>% 
  filter(location != "Nomad")

center <- list(
  "Unweighted Average" = geographic_average(
    lon = df$lon,
    lat = df$lat),
  "Weighted Average" = geographic_average(
    lon = df$lon,
    lat = df$lat,
    weight = df$number),
  "Unweighted Midpoint" = geographic_midpoint(
    lon = df$lon,
    lat = df$lat),
  "Weighted Midpoint" = geographic_midpoint(
    lon = df$lon, 
    lat = df$lat,
    weight = df$number)
)

center <- bind_rows(center, .id = "calculation")
```

```{r echo = FALSE}
leaflet() %>%
  addTiles() %>%  
  addCircleMarkers(lng = center$lon,
                   lat = center$lat,
                   label = center$calculation,
                   radius = 8)
```

We can see that the weighted and unweighted geographic midpoints lie in Northeastern middle-of-nowhere Canada, and the weighted and unweighted averages lie in the Atlantic Ocean off the coast of the United States.

We can calculate the difference in the total distance that the team needs to travel to get to this midpoint. First, let's calculate the total distance traveled to get to San Diego.

```{r}
# set location
locations$destination <- "San Diego, CA, USA"

# get destination coordinates
dest_coords <- geocode_locations("San Diego, CA, USA")
target_coords <- c(32.71725, -117.17)

# set coordinates
locations$lat2 <- 32.71725
locations$lon2 <- -117.17

locations$distance <- distHaversine(locations[,4:3], locations[,7:6]) / 1000

locations %>% 
  filter(location != "Nomad") %>% 
  summarise(kms = sum(distance))
```

425 thousand kilometers! That's around 10 times around the Earth! Let's calculate the total distance to the weighted midpoint, which appears to be in the middle of nowhere.

```{r}
# set coordinates
locations$lat2 <- 55.37969
locations$lon2 <- -73.73865

locations$distance <- distHaversine(locations[,4:3], locations[,7:6]) / 1000

locations %>% 
  filter(location != "Nomad") %>% 
  summarise(kms = sum(distance))
```

The team would only have to travel 343 thousand kilometers to get to the geographic midpoint. That's around 20% less kilometers traveled and only 8.5 times around the world. Let's calculate the distance to the weighted average.

```{r}
# set coordinates
locations$lat2 <- 39.27205
locations$lon2 <- -56.07747

locations$distance <- distHaversine(locations[,4:3], locations[,7:6]) / 1000

locations %>% 
  filter(location != "Nomad") %>% 
  summarise(kms = sum(distance))
```

It would be around 398 thousand kilometers to the geographic average in the Atlantic Ocean.
